---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---




```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set( results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, fig.dim = c(7,7), autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}

### Load general Deps
cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_variableGenes.R"));
read_chunk( path=file.path( WORKING_DIR, "03_cellHeterogeneity.R"));
read_chunk( path=file.path( WORKING_DIR, "04_characterizeCluster.R"));
```

---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
# outputFilesPrefix = paste0( SCIENTIFIC_PROJECT_NAME, "_",     
#                             EXPERIMENT_PROJECT_NAME, "_"
#                             #startTimeFileName, "_",
#                             #paramsHash, "_"
#                             )

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
main_heatmap(matrix(0));
cat('</div>');
```


## Heterogeneity analysis on all samples


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = SAMPLE_NAME,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />




### Load data from QC analysis and remove contamination

In this section, we load the data filtered after QC and we remove the cells identified as contamination.

```{r dataLoad}
<<loadData>>
```
<br /><br />

### Variable genes
```{r findVariableGenes}
<<findVariableGenes_seuratMethod>>
```

Number of variable genes found: `r variablesGenesStats`

```{r findVariableGenes_summaryPlot}
<<findVariableGenes_summaryPlot>>
```
<br />

Top `r VARIABLE_FEATURES_SHOWTOP` most variable annotations:

```{r findVariableGenes_summaryTable}
<<findVariableGenes_summaryTable>>
```
<br /><br />



### Clusters identification

#### Cells distribution
```{r heterogeneity_identifyClusters}
<<heterogeneity_identifyClusters>>
```

#### Statistics by cluster
```{r heterogeneity_identifyClusters_splitStats}
<<heterogeneity_identifyClusters_splitStats>>
```
<br /><br />


### PCA results

PCA projections for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca>>
```
<br />

UMIs counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_umisCounts, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca_umisCounts>>
```
<br />

Genes counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_genesCounts, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca_genesCounts>>
```
<br />

Correlation (Spearman) of PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")` with UMIs and genes counts  
```{r heterogeneity_pca_correlations, fig.dim = c(9,6)}
<<heterogeneity_pca_correlations>>
```
<br />

PCA loadings of top `r PCA_PLOTS_NBFEATURES` features for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_loadings, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca_loadings>>
```
<br /><br />




### Dimensionality reduction and clustering {.tabset .tabset-fade}
```{r heterogeneity_dimReduc}
<<heterogeneity_dimReduc>>
```

#### tSNE {.tabset .tabset-fade}
```{r heterogeneity_dimReduc_plot_tSNE}
useReduction='tsne';

<<heterogeneity_dimReduc_interactivePlot_clusters>>
#<<heterogeneity_dimReduc_interactivePlot_samples>>
#<<heterogeneity_dimReduc_details_samples>>
rm("useReduction");
```

#### UMAP {.tabset .tabset-fade}
```{r heterogeneity_dimReduc_plot_UMAP}
useReduction='umap'

<<heterogeneity_dimReduc_interactivePlot_clusters>>
#<<heterogeneity_dimReduc_interactivePlot_samples>>
#<<heterogeneity_dimReduc_details_samples>>
rm("useReduction");
```


<!-- Ending the tabset section -->
#### {.toc-ignore}

<!-- Ending the tabset section -->
### {.toc-ignore}
<br /><br />




### Marker genes
```{r heterogeneity_markerGenes}
<<heterogeneity_markerGenes>>
```

#### Marker genes list
```{r heterogeneity_markerGenes_table}
<<heterogeneity_markerGenes_table>>
```
<br />

#### Marker genes heatmap

Expression heatmap of top marker genes sorted by Seurat clusters:

```{r heterogeneity_markerGenes_heatmap}
<<heterogeneity_markerGenes_heatmap>>
```

#### Marker genes expression on t-SNE/UMAP {.tabset .tabset-fade}

##### tSNE {.tabset .tabset-fade}
```{r heterogeneity_markerGenes_expression_tSNE, fig.dim = c(3,3), out.width='25%'}
useReduction='tsne';

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

##### UMAP {.tabset .tabset-fade}
```{r heterogeneity_markerGenes_expression_UMAP, fig.dim = c(3,3), out.width='25%'}
useReduction='umap';

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}


<!-- Ending the tabset section -->
#### {.toc-ignore}
<br /><br />


#### Marker genes expression on violin plots

Here we show the detail of expression of all genes identified as marker for at least one cluster
as violin plot per cluster (with jitter point). Numbers above each violin plot is (i) top: the number
of cells with a positive expression of the gene in the cluster (ii) bottom : the number of cells
with a null expression of the gene in the cluster.

```{r heterogeneity_markerGenes_expression_violin, fig.keep = "all", fig.dim = c(3,3), out.width='25%'}
<<heterogeneity_markerGenes_expression_violinplot>>
```

### Monitored genes
```{r heterogeneity_monitoredGenes}
<<heterogeneity_monitoredGenes>>
```

#### Monitored genes heatmap

Expression heatmap of monitored genes sorted by Seurat clusters:

```{r heterogeneity_monitoredGenes_heatmap}
<<heterogeneity_monitoredGenes_heatmap>>
```

#### Monitored genes expression {.tabset .tabset-fade}

##### tSNE {.tabset .tabset-fade}
```{r heterogeneity_monitoredGenes_expression_tSNE, fig.dim = c(3,3), out.width='25%'}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

##### UMAP {.tabset .tabset-fade}
```{r heterogeneity_monitoredGenes_expression_UMAP, fig.dim = c(3,3), out.width='25%'}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

<!-- Ending the tabset section -->
#### {.toc-ignore}
<br /><br />

#### Monitored genes expression on violin plots

Here we show the detail of expression of all monitored genes
as violin plot per cluster (with jitter point). Numbers above each violin plot is (i) top: the number
of cells with a positive expression of the gene in the cluster (ii) bottom : the number of cells
with a null expression of the gene in the cluster.

```{r heterogeneity_monitoredGenes_expression_violin, fig.keep = "all", fig.dim = c(3,3), out.width='25%'}
<<heterogeneity_monitoredGenes_expression_violinplot>>
```

<!--

### Characterizing clusters

#### Looking at clusters similarity by comparing pairwise marker genes

In this section we compute the pairwise gene markers between each two clusters. Then we look at the similarity bwteen clusters. This similarity is computed by comparing (i) the number A of genes that are over-expressed in the first cluster resepetc to the second one and (ii) the total number B of pairwise marker genes between all pairs of clusters.
The similarity is equal to 100*( 1 - (A/B)). 

So a high similarity means that the the cluster do not have a lot of over-expressed genes respect to the second cluster regarding the global over-expression among all the clusters. So, the two clusters are relatively similar regarind the global heterogeneity in the sample.

```{r heterogeneity_pariwiseMarkerGenes}
#<<characterizeCluster_pairwiseMarkers>>
#<<characterizeCluster_similarity>>
```

#### Looking at enriched biological processes by pairwise gene markers

In this analysis we aim to characterize the clusters biological functions. We consider we wish to know only the difference between the clusters, i.e. understand through all the biological processes activated in all the clusters which one are activated by which cluster and if this biological process is characteristic of the cluster respect to the others.

To do so, we first execute a pairwise comparison of clusters to extract for each of them the list of positive marker genes : for each cluster, we make (n-1) comparison (n being the number of cluster), each comparison providing a list of genes over-expressed in the first cluster compared to the second.

Next, for each cluster, we look a its functionnal enrichments : we consider the union of all the pairwise marker genes of this cluster against the GO biological process genesets with the union of all the pairwise marker genes of all clusters as background.

This analysis may show the biologial processes (BP) that are specifically activated by a cluster respect to all the other clusters. For instance, if a BP is activated by only one or two clusters among 10, this BP may be shown as significantly enriched in those two clusters. At the contrary, if a BP is activated by almost all clusters, it may not be shown as significant for any of the clusters.

This analysis is relative to the clusters since it's goal is to relatively characterize each cluster respect to the others.

```{r heterogeneity_pariwiseMarkerGenes_functionalEnrichment, warning=FALSE, message=FALSE}
#<<characterizeCluster_functionalEnrichment>>
```

#### Looking at enriched biological processes by pairwise gene markers looking only at neighbors genes

In this section we do tthe same analysis as before bu the background used for the enrichment analysis is limited to the cluster that are the neighbors of the targeted cluster.

```{r heterogeneity_pariwiseMarkerGenes_functionalEnrichment_fromNeighbors, warning=FALSE, message=FALSE}
#<<characterizeCluster_functionalEnrichment_fromNeighbors>>
```
-->

### Exporting results to file

```{r exportResults, results='asis', echo=FALSE, warning=FALSE}

cat("<BR>Exporting expression data to file:<BR>")

# Exporting raw expression data (as produced by Cellranger)
raw_expression_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Contaminationfiltered_raw_expression_matrix.csv"))
cat("<BR>* Raw expression:<BR>", raw_expression_file_path)
write.table( as.data.frame( GetAssayData(object = sc10x.rna.seurat, slot = "counts")), file = raw_expression_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

# Exporting normalized expression data
normalized_expression_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Contaminationfiltered_expression_matrix.csv"))
cat("<BR>* Normalized expression:<BR>", normalized_expression_file_path)
write.table( as.data.frame( GetAssayData(object = sc10x.rna.seurat, slot = "data")), file = normalized_expression_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

cat("<BR><BR>Exporting Rdata to files:<BR>")

# Exporting Seurat object
seurat_object_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Contaminationfiltered_sc10x.rna.seurat.RDS"))

cat("<BR>* Seurat Object:<BR>", seurat_object_file_path)
saveRDS( object = sc10x.rna.seurat, file = seurat_object_file_path)

cat("<BR><BR>Exporting dimensionality reduction mapping to files:<BR>")

# Exporting PCA map
PCA_map_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Contaminationfiltered_PCAcoord.tsv"))
cat("<BR>* PCA map:<BR>", PCA_map_file_path)
write.table( Embeddings( sc10x.rna.seurat, reduction = "pca"), file = PCA_map_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

# Exporting t-SNE map
tsne_map_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Contaminationfiltered_Tsnecoord.tsv"))
cat("<BR>* t-SNE map:<BR>", tsne_map_file_path)
write.table( Embeddings( sc10x.rna.seurat, reduction = "tsne"), file = tsne_map_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

# Exporting UMAP map
umap_map_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Contaminationfiltered_Umapcoord.tsv"))
cat("<BR>* UMAP map:<BR>", umap_map_file_path)
write.table( Embeddings( sc10x.rna.seurat, reduction = "umap"), file = umap_map_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

cat("<BR><BR>Exporting clustering to files:<BR>")
seurat_clusters_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Contaminationfiltered_Seurat_clusters.tsv"))
cat("<BR>Seurat Clusters:<BR>", seurat_clusters_file_path)
Idents( sc10x.rna.seurat) = "seurat_clusters"
write.table( data.frame( Idents( sc10x.rna.seurat)), file = seurat_clusters_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")


```


