---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---




```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set( results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, fig.dim = c(7,7), autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}

### Load general Deps
cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"))
read_chunk( path=file.path( WORKING_DIR, "02_velocityAnalysis.R"))
```

---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
# outputFilesPrefix = paste0( SCIENTIFIC_PROJECT_NAME, "_",     
#                             EXPERIMENT_PROJECT_NAME, "_"
#                             #startTimeFileName, "_",
#                             #paramsHash, "_"
#                             )

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
main_heatmap(matrix(0));
cat('</div>');
```




## Heterogeneity analysis on all samples


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = SAMPLE_NAME,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />


# Single-cell Analysis with Velocyto

## Raw Analysis

#### Loading and building data
In this section, we load the data, filter the cells using the previous analysis (QC and contamiantion) and load the various embedding (t-SNE, UMAP) and the cluster computation from
the previous analysis
```{r loadingData, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
<<load_data>>
<<filter_data>>
```  

### Velocity estimation
In this section, we look for the velocity of cells mapped on embeddings (t-SNE, UMAP) using or not the grid smoothing

```{r estimateVelocity, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
<<estimate_velocity>>
```  
