---
title: "SPlab - ILC you project Secondary Analysis"
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    # dev: "pdf"
    # self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

```{r loadLibraries, echo=FALSE, warning=FALSE, message=FALSE}
library( DESeq2)
library( ggplot2)
library( ggrepel)
library( xtable)
library( stringr)
library( biomaRt)
library( ade4)
library( mgcv)
library( GGally)
library( pheatmap)
library( RColorBrewer)
library( gage)
library( org.Mm.eg.db)
library( knitr)
library( kableExtra)
library( DT)
library( d3heatmap)
library( htmltools)
library( ggiraph)
library( reshape)
```

```{r definePathConstants, echo=FALSE, warning=FALSE, message=FALSE}


dir.create( OUTPUT_DIR, showWarnings = FALSE, recursive = TRUE)
```

```{r defineAnalysisConstants, echo=FALSE, warning=FALSE, message=FALSE}
POPULATION_TYPE = list( EP= c("EP1", "EP2", "EP3", "EP4"), FLP= c("FLP1", "FLP2", "FLP3", "FLP4"))
UNWANTED_REPLICATES = c( "FLP4.2")

MONITORED_GENES = c( "Flt3", "Ets1", "Lta", "CD4", "Rorc", "Il17a", "Il23r", "Il22", "Itgb5", "Zbtb16",
                           "Itga4", "Itgb7", "Tnfsf11", "Tnfrsf11a", "Id2", "Cxcr4", "Cxcr5", "Nfil3", "Tcf7", "Tox")

GENE_FUNCTIONAL_CATEGORIES = list(
             Transcription_factors = c( "Rora", "Rorc", "Ahr", "Arnt2", "Gata3", "Icos", "Id1", "Id2", "Id3", 
                                        "Nfil3", "Runx1", "Tcf7", "Tox", "Tox2", "Zbtb16"),
             LTi_markers = c( "Itga4", "Itgb7", "Itgb5", "Cxcr6", "Ccr6",  "S1pr1", "Cd34", "Cd4", "Lta", "Ltb"),
             Cytokines = c( "Il17a", "Il17f", "Il17ra", "Il17re", "Il22", "Il23r", "Il27ra", "Il2rb"),
             Notch_signaling = c( "Dll1", "Hes1", "Batf3", "Cd160", "Cd27", "Ikzf3", "Arg1"),
             TNF_related = c( "Tnfrsf13b", "Tnfrsf11a", "Tnfrsf25", "Tnfrsf9", "Tnfsf9"),
             Retinoic_acid = c( "Fabp4", "Fabp7", "Rxrg", "Rbp4", "Stra6", "Rarb", "Rara", "Rarg", "Rxra", "Rxrb")
)
```

```{r definePlotConstants, echo=FALSE, warning=FALSE, message=FALSE}
AXIS_TICK_SIZE = 12
AXIS_TITLE_SIZE = 14
LEGEND_TEXT_SIZE = 14

POPULATION_NAME_MAPPING = c( EP1 = "eCLP", EP2 = "eLTiP", EP3 = "eLTi0", EP4 = "eLTi4",
                             FLP1 = "fCLP", FLP2 = "fLTiP", FLP3 = "fLTi0", FLP4 = "fLTi4")
```


[//]: # "Load the various chunks in the various files"
```{r readChunks, echo=FALSE, warning=FALSE, message=FALSE}
read_chunk( path=file.path( SCRIPT_DIR, "prepare_load_data.R"))
read_chunk( path=file.path( SCRIPT_DIR, "analysis_expression_pattern.R"))
```

## Read the data

```{r loadData, echo=FALSE, message=FALSE, results = "asis"}
<<load_data>>
```  

## Evolution pattern along EP populations

In this section, we determine the evolution pattern of each genes along the four EP populations. Evolution
are encoded with 3 numbers for the variation between EP1-EP2, EP2-EP3 and EP3-EP4. a "1" indicates an increase of value, a "-1" a decrease and a "0" no variation.

For instance:
* "111" = increase between EP1 and EP2, between EP2 and EP3 and between EP3 and EP4
* "1-11" = increase between EP1 and EP2, decrease between EP2 and EP3 and increase between EP3 and EP4

Note : We show only the top 20 gene with the most important variations between populations.

```{r expressionPattern, echo=FALSE, message=FALSE, results = "asis"}
<<expression_pattern>>
```  

